Homicide Increases in the COVID-19 Era

### 2025-02-27 Draft1.qmd
updated
regular save

crime_la_monthly$high_low_crime <- as.factor(ifelse(crime_la_monthly$num_crimes > avg_crimes, "Above Average", "Below Average"))

@@

```{r}
crime_by_area <- crime_la %>% group_by(AREA.NAME) %>% summarize(numberOfCrimes = n()) %>% ungroup()

@@

```{r, echo=FALSE}
# creating dataset for correlations

# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
```

>>>>

@@

```{r, echo=FALSE}
crime_by_area <- crime_la %>% group_by(AREA.NAME) %>% summarize(numberOfCrimes = n()) %>% ungroup()

@@

```{r, echo=FALSE}
library(lubridate)
# creating dataset for correlations

# since we do not have pre- and post- pandemic city of LA data, we are instead using LA county data for evictions. 
LA_county_evictions <- read.csv("Raw_Data/LA_county_evictions.csv")

# covid data
covid_la$date <- ymd(covid_la$ep_date)
covid_la$month <- month(covid_la$date)
covid_la$year <- year(covid_la$date)
covid_la$date <- day(covid_la$date)
# since there is a 7 day lag in reporting, let's do monthly = cases on the 21st and on the 5th of the next month.
covid_la$lagged <- covid_la$date + 7

covid_la_lagged <- covid_la %>% filter(lagged %in% c(21, 35)) %>% arrange(month, year)

covid_la_monthly <- covid_la_lagged %>% group_by(month, year) %>% summarize(cases = sum(cases_14day)) %>% ungroup() %>% arrange(year, month)

covid_la_monthly$index <- 1:nrow(covid_la_monthly)
# base plot shows that this does in fact mirror LA county so the case numbers should be accurate
# plot(y = covid_la_monthly$cases, x = covid_la_monthly$index, type = "l")

crime_la_monthly$date <- ymd(paste0(crime_la_monthly$month_year, "-01"))
crime_la_monthly$month <- month(crime_la_monthly$date)
crime_la_monthly$year <- year(crime_la_monthly$date)
```
